---
title: "Bumble bee competition revisited under global climate change"
author:
- Douglas B. Sponsler
- Fabrice Requier
- Katharina Kallnik
- Alice Claßen
- A. Fabienne Maihoff
- Ingolf Steffan-Dewenter 
output:
  pdf_document: default
  word_document:
    reference_docx: docx_template.docx
bibliography: refs.bib
---

```{r echo = FALSE, include = FALSE}
library(tidyverse)
```

**Introduction**

The nascent reality of climate change has already revealed the fragile contingency of ecological paradigms that rely on the assumption of climate stability [@Hannah2002-nv; @Walther2002-mq]. Thus, one of the central questions facing ecologists today is how climate change will affect the future fitness and distribution of species and, thereby, the composition and function of ecological communities. While it is intuitive that climate change should affect species and communities through environmental filtering, it is now recognized that climate change can also restructure biotic interactions [@Tylianakis2008-fi], such as competition and facilitation [@Ettinger2017-pk], altering the fitness landscape that governs community assembly and species coexistence [@Chesson2000-aw].  

Bumble bees (Hymenoptera: *Bombus* spp.) have become a model system for the study of biotic interactions in ecological communities. In particular, the broad and overlapping distributions of bumble bee species, together with their gross similarity in morphology, behavior, and dietary niche, have attracted the attention of researchers interested in questions of interspecific competition and, more broadly, the rules of community assembly and species coexistence [@Plowright1984-lc; @Goulson2008-zy]. In this regard, however, bumble bees have proven somewhat enigmatic. On the one hand, there is evidence that resource limitation can be a significant constraint on local bumble populations [@Heinrich1976-wc; @Bowers1985-df], and certain bumble bee communities have been shown to be over-dispersed in terms of tongue length (a matching trait corresponding to floral corolla depth) or phyloogeny (understood to encompass unmeasured trait variation), providing evidence for competition as a community assembly process [e.g. @Inouye1977-kl; @Pellissier2013-pk]. Nevertheless, bumble bee communities with an apparent excess of species richness and deficit of trait (or phylogenetic) dispersion are commonly observed [@Ranta1981-vx]. To account for this inconsistency, @Ranta1981-vx proposed that bumble bees communities are governed by a spatiotemporally heterogeneous mosaic of competitive conditions, such that competitive inequalities between species are sufficiently variable to allow more species to coexist than predicted by trait-based competition models. A complementary explanation, based on the core/satellite species hypothesis [@Hanski1982-hs], is that competitive forces are significant only among regionally abundant and widespread "core species", and the effects of competition are more pronounced in the structure of regional species pools than in that of local assemblages [@Hanski1982-xd]. 

As pointed out by @Williams1989-vr, however, the question of competition as a biogeographic/evolutionary-scale driver of community assembly differs from that of competition as a local/behavioral-scale driver of species' interactions. While the role of competition in bumble bee community assembly remains somewhat obscure, there is broad empirical evidence that bumble bees facultatively partition resources to minimize competition [e.g @Inouye1978-qi; @Morse1982-bh; @Ishii2013-sm; @Brosi2013-mu; @Brosi2017-oc]. This distinction between spatiotemporal scales of competition, echoing the logic of @Ranta1981-vx, becomes especially important when the scope of inquiry switches from retrospective to prospective, and all the more so when there is reason to think that the ecological context of competition is undergoing rapid and unprecedented change. Since the height of bumble bee competition research in the 1980s, such a shift has indeed occurred, with retrospective questions of community assembly all but eclipsed by the urgency of conservation in the face of global changes in climate and land use [@Goulson2010-lo]. Given the uncertain spatiotemporal lags and exogenous drivers of community assembly, inferring the past role of competition through the study of bumble bee community structure may be of limited use in predicting the future fitness and distribution of bumble bee species. In this regard, the study of competition via behavioral patterns of floral resource partitioning that respond in real-time to variation in ecological context [@Spiesman2016-nf] provides a vital complement to the classical study of community assembly.  

The study of competition as a behavioral phenomenon, however, is empirically and analytically demanding, since species interactions, not merely community composition, must be measured and interpreted. The difficulty is compounded when the goal is to infer the effects of climate variation, since observations must be replicated across gradients of interest while minimizing the influence of confounding spatial or temporal effects. An elegant solution to both these challenges is the use of species interaction network analysis along steep elevation gradients [e.g. @Hoiss2015-gv; @Albrecht2018-gn]. Interaction network analysis provides a tractable framework for the study of resource partitioning [@Bluthgen2006-if], and steep elevation gradients can serve as space-for-time proxies in the study of climate effects, capturing a broad range of climate variation while minimizing the confounding effects of geographic distance [@Blois2013-wr]. 

Previous studies of pollinator and plant communities along elevation gradients have tended to find decreased specialization and/or increased niche overlap at high elevation, which is often interpreted as an indication of reduced interspecific competition, though other mechanisms cannot be excluded [@Hoiss2012-sn; @Miller-Struttmann2014-wz; @Hoiss2015-gv; @Classen2020-nh]. If, in fact, competition in pollinator communities is attenuated at high elevation, this would be broadly consistent with the "stress-gradient hypothesis", which posits that prevailing biotic interactions shift from negative (e.g. competition) to positive (e.g. facilitation) along ascending gradients of environmental stress [@Bertness1994-zx]. As applied to pollinator communities, the logic would be that environmental stresses at high elevation, such as low temperatures and short growing seasons, suppress pollinator densities more strongly than the densities of their shared resources, thus causing a per-capita increase in food availability and a corresponding relaxation of competition. 

In this study, we analyze patterns of resource partitioning in a bumble- bee-wildflower network sampled along an elevation gradient in the Berchtesgadener Alps of southeastern Germany. We first analyze species- and group-level network metrics of specialization and niche overlap to determine whether the bumble bees in our study system exihibit the expected pattern of decreased specialization and increased niche overlap at high elevation. To evaluate the plausibility of intepreting specialization and niche overlap as indices of competition, we then use independent floral surveying data to determine whether observed patterns in specialization and niche overlap correspond to variation in per-capita resource availability. Finally, we use the within-year temporal resolution of our dataset to test for reciprocal population dynamics of species within sites as a direct indication of interspecific competition. From these results, we derive predictions concerning the effects of climate change on competition as a constraint on bumble bee species and communities.  



**Methods**\

*Field sampling.*  Our study is set in the Berchtesgadener Alps of southwestern Germany, a system in which both plant and pollinator communities have been well-studied [@Hoiss2012-sn; Hoiss2012-kr; Hoiss2015-gv]. Our study sites are the same as those  ranging in elevation from 641 to 2032 m above sea level. 

*Degree day modeling.*

*Colony life cycle analysis.* Bumble bee colonies exhibit three distinct life history phases: a "founding" phase during which colonies are initiated by solitary queens, a "build-up" phase during which workers are produced, and a "reproductive" phase where colonies produce new queens and males [@Goulson2010-lo]. A formal delimitation of these phases serves both as a temporal framework for analyzing bumble bee and floral dynamics and as a check that different elevational subsets of our data capture the same life history interval and are thus comparable in terms of network structure.

To delimit colony phases, we fit binomial regression models to the proportional abundance of foundress queens (Q) and new reproductives (R) (males + new queens), respectively. We set the breakpoint between the founding and build-up stages at Q = 0.75, and the breakpoint between buildup and reproductive stages at R = 0.1. These cutoffs are somewhat arbitrary points in a continuous process, but their primary purpose is to facilitate comparison across sites. This analysis was performed on both a GDD and Julian temporal scale. All operations were done in R. 


*Inferring resource limitation.* During the founding phase, it would be hard to detect resource limitation even if it were occurring, since the population would not be able to respond to it except by actual starvation of queens. In other words, until workers emerge, the relationship between bumble bee density and floral density would be expected to be null. At least if competition is the proposed mechanism. If mere concentration of foragers is the mechanism, than you would expect to see BB density respond positively to floral density, and the fact that it does not in our data suggests that, at least during the founding phase, forager concentration is not an important mechanism. That's good news.


**Results**\
*Colony life cycle analysis.*  
```{r, echo = FALSE, message = FALSE, fig.cap = "**Figure 2:** Colony life cycle phases delimited by the proportional abundances of foundress queens and new reproductives. Dashed lines indicate where foundress queen and reproductive proportional abundances cross the thresholds of 0.75 and 0.10, respectively. Binomial regression curves are fit to the proportional abundances, with a spline term added to capture the unimodal pattern of worker abundance (not used in calculating breakpoints)."}
breaks_yday_summary <- read_csv("./output/breaks_yday.csv") %>%
  mutate(elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))
breaks_gdd_summary <- read_csv("./output/breaks_gdd.csv") %>%
  mutate(elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))

# Binomial ggplot smoother
binomial_smooth <- function(...) {
  geom_smooth(method = "glm", method.args = list(family = "quasibinomial"), se = FALSE, ...)
}

ggplot(breaks_yday_summary, aes(yday, prop, color = caste)) +
  geom_point() +
  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
  geom_vline(aes(xintercept = break1), linetype = "dashed") +
  geom_vline(aes(xintercept = break2), linetype = "dashed") +
  facet_grid(elev.class2 ~ year) +
  xlab("Julian day") +
  ylab("Proportional abundance")

# ggplot(breaks_gdd_summary, aes(gdd.cum, prop, color = caste)) +
#   geom_point() +
#   binomial_smooth(formula = y ~ splines::ns(x, 2)) +
#   geom_vline(aes(xintercept = break1), linetype = "dashed") +
#   geom_vline(aes(xintercept = break2), linetype = "dashed") +
#   facet_grid(elev.class2 ~ year) +
#   xlab("GDD") +
#   ylab("Proportional abundance")
```
The colony life cycle intervals captured by our sampling varied across years, but were consistent across elevational classes within years, validating the comparison of foraging patterns along our elevation gradient **(Figure 2)**. In 2010, all three colony phases are represented in our samples, but the founding phase appears somewhat truncated, suggesting that sampling began after colony founding was already underway. In 2011, the founding phase was missed entirely, with sampling beginning mid-way through the build-up phase. In 2012, we achieved the longest and densest sampling, yielding good representation of all three colony phases. It is worth noting that as elevation increases, colony lifecycle is delayed in onset and contracts in duration, but the phase breakpoints remain approximately equal (on a Julian time scale) across elevation classes.   

*Specialization and niche overlap*. 

*Resource limitation*.   Per-transect bumble bee abundance is positively correlated with per-transect floral cover (log-transformed). To the extent, therefore, that our metrics of bumble bee and floral density reflect true site-scale abundance and not just co-aggregation, this pattern suggests that bumble bee
density is resource limited, particularly at sites with very low floral density. Importantly, this approach does not distinguish between inter- and intraspecies (or even intra-colony) competition [@Ranta1981-vx].

Note, moreover, that this relationship varies seasonally. For now, we will use Julian month as a stand-in for GDD, which we'll use once we have analyzed our weather station data. Early in the year, the relationship between BB density and floral density is weak; this makes sense because colonies are just getting established during the first couple of months, and BB density is more limited by the autocorrelation component of colony size (bigger colonies can raise more brood) than by floral resource availability. In mid- to late-season, however, BB density becomes strongly dependent on floral resource density. Interestingly, in September, the relationship is strongest and floral density is also lowest.

*Reciprocal population dynamics*. 

*Colony life cycle analysis.* 

**Discussion**\




<!-- *Study sites.*  -->

<!-- *Field sampling.*  -->

<!-- *Data analysis.*  -->
<!-- \ -->


<!-- **Results** -->


<!-- So, we have at least circumstantial evidence that bumble bee density is limited by floral density (again, the key assumption is that our -->
<!-- metrics of bumble bee and floral density reflect true site-scale abundance). What we might as next is whether the degree of dependence of bumble bee density upon floral density vary with elevation. It's hard to draw conclusions about the lowest elevation class, but it seems pretty clear that the slope of bumble bee density on floral density is steep at the highest elevation class than at the middle elevation class; moreover, the relationship is tighter (lower variance) at high elevation. It appears that, contrary to the predictions of the Ranta hypothesis, bumble bees are more dependent on floral resource availability at higher elevations. -->

<!-- But let's look at this another way. By dividing bumble bee density by floral density, we can derive an index that represents "bees per unit of floral abundance" (which we willlog-transform for this analysis to correct for skewed distribution). We'll call this our "bee-packing" index, and we'll understand it as an estimate of predicted competition intensity. When we plot this against elevation, parsed by year, we see that "bee-packing" seems to increase at high elevation for 2011, and possibly also in 2010. But the relationship seems to disappear in 2012. This is worth thinking about. 2012 was a year of very low floral density. Perhaps this increased resource limitation across the whole elevation gradient, erasing the effect of elevation. -->

<!-- Some pretty neat patterns emerge when we break our data down by month. For most months, bee-packing increases with elevation, though sometimes the relationship is fairly flat. In Septmeber, however, bee-packing decreases with elevation, presumably due to the elevational staggering of phenology; by that point in the year, only the colder sites still have appreciable floral density. -->

<!-- Another way to probe the question of competition is to examine network structure, though we will have to sacrifice our temporal resolution to do so. An advantage of this analysis is that we will no longer rely on the estimate of floral density using our survey data, so we can shed the somewhat vulnerable assumption that the correlation between bumble bee density and floral density is not due merely to co-aggregation. To do this, we will focus on the group-level metrics of niche overlap and C-score. Niche overlap would be expected to vary inversely with competition, while C-score -- a measure of ... -- would be expected to vary directly with competition. For this analysis to be sound, though, we need to rarefy our networks. Otherwise, niche overlap and C-score could be driven by uneven sampling effort (less intensively sampled sites will appear to have lower niche overlap and higher C-score). -->


<!-- **Discussion** -->

<!-- -  Ingolf said that he does not think that competition from syrphids at high elevation would be significant, but competition from honey bees at low elevation could be -->

<!-- Bumble bee communities are structured by a spatiotemporal mosaic of competitive -->
<!--   processes -->

<!-- The two processes of environmental filtering and biotic interactions may be difficult to disentangle [@Kraft2015-gh].  -->


<!-- The nascent reality of climate change, and the grave expectation of its progression in the coming decades, have revealed the fragile contingency of conservation paradigms that rely on the implicit assumption of climatic stability [@Hannah2002-nv]. -->

<!-- Under conditions of resource limitation, intraspecific competition may be a stronger constraint on bumble bee density (though not on diversity) than interspecific competition, since conspecific are limited in their ability to partition further their potential niche space. -->

<!-- "Altitudinal niche-breadth hypothesis" [@Rasmann2014-zb]: Analogous to the latitudinal niche-breadth hypothesis, organisms at high elevations tend to have broader niches due to higher environmental variability, and possibly other reasons. Interestingly, @Rasmann2014-zb do not invoke the concept of competition release as a reason for increased niche breadth at higher elevations. -->

<!-- It is important to clarify what we are and are not sampling. We are not sampling bumble bee communities; we are sampling spatial (and temporal) subsets of the floral resource landscape. So, the questions we are able to answer concern not the composition of the community as such, however one chooses to delimit it, but rather the composition of foragers within an arbitrarily delimited (but standardized) segment of the resource landscape (i.e. the sampling plot). -->

<!-- Importantly, though, competition need not be universal, or even frequent, to be a powerful driver of community assembly [@MacArthur1984-qz]. -->

<!-- Moreover, trait matching under interspecific competition might also exceed that predicted under no competition, since interspecific competition would limit a species' ability to "adaptively mismatch" to exploit a rich but poorly matched resource, since the resource would likely be monopolized by a better-matched (and therefore more competitive) species. -->

<!-- Am I right in thinking that it is impossible to have interspecific competition without also having intraspecific competition? These processes would have opposite effects on niche breadth. By itself, interspecific competition would favor the maximal contraction of niche breadth until no overlap remained. But intraspecific competition would have the opposite effect, forcing conspecifics to broaden their niches to minimize competition with one another. Presumably, observed niche breadth and niche overlap represent a dynamic equilibrium of these opposing forces. -->



<!-- Our analysis has a hierarchical structure. First, we test the hypothesis that competition is of negligible significance in our study system. At least two plausible mechanisms can be proposed that would account for a lack of competition in bumble bee communities [@Ranta1981-vx]. First, it is possible that a super-abundance of floral resources could saturate the foraging capacity of a local bumble bee population. Indeed, this may be a common (though transient) phenomenon associated with the bloom of mass flowering crops, such as red clover [@Ranta1981-vx] and oilseed rape [@Westphal2009-sb]. Alternatively, competition could be attenuated if abiotic stress, such as low temperature, suppresses bumble bee populations below competition-inducing densities. This pattern would be a prediction of the "stress-gradient hypothesis" [@Bertness1994-zx], and while it has not, to our knowledge, been documented for bumble bees specifically, a recent study of whole bee communities in the Berchtesgaden system (of which bumble bees were a part) found evidence of dampened competition at high elevation sites along a montane-to-alpine gradient [@Hoiss2012-sn]. Under either mechanism, the most straightforward prediction is that bumble bee population density would be decoupled from resource density, since the latter would be non-limiting. For the same reason, the absence of competition would also predict that resource partitioning would be constrained only by trait-matching, not by niche overlap; thus, niche partitioning among species with similar traits should not exceed that predicted by a random null model.  -->

<!-- Where the hypothesis of no competition can be rejected, we seek to differentiate inter- from intraspecific competition. These two forms of competition share the necessary precondition of resource limitation, but they imply distinct predictions regarding resource partitioning and trait-matching. Resource partitioning by bumble bees is mediated by variation in matching traits, the most well-studied of these being tongue length. Bumble bee tongue lengths vary between species, and the efficiency with which a bumble bee can collect nectar from a given flower depends largely on the closeness of match between the tongue length of the bee and the corolla depth of the flower [@Inouye1980-mv; @Ranta1980-ok]. Since tongue length also varies within species, and literature values for a given species vary, it is common to average available values and bin species into tongue-length "guilds" [@Pellissier2013-pk]. -->

<!-- Where interspecific competition is strong, a species' dietary specialization should increase in proportion to the density of other species within the same tongue length guild per density of floral resources shared by that guild (hereafter referred to as "guild-packing"). Moreover, species within a tongue length guild should tend to exhibit divergent abundance dynamics, since they are reciprocally limiting. Under the pressure of intraspecific competition, however, individuals may respond by foraging more opportunistically within the constraints of their fundamental niche space []. Alternatively, a species may relieve intraspecific competition by subdividing, on the basis of intraspecific trait heterogeneity, into differently specialized populations [@Araujo2008-bp]. In either case, the species-level effect would be an increase in realized niche breadth (i.e. a decrease in specialization) in proportion to conspecific-packing (density of conspecifics per density of guild-level floral resources) [though see @Parent2014-hx].  -->

<!-- These hypotheses and predictions are summarized in Table 1. -->

<!-- +--------------------------------+--------------------------------------------------------+ -->
<!-- | Hypothesis                     | Prediction                                             | -->
<!-- +================================+========================================================+ -->
<!-- | H1: No competition             | 1. BB density !~ FL density                            | -->
<!-- +--------------------------------+--------------------------------------------------------+ -->
<!-- | H2a: Interspecific competition | 1. specialization ~ guild-packing                      | -->
<!-- |                                | 2. within-guild abundance divergence                   |  -->
<!-- +--------------------------------+--------------------------------------------------------+ -->
<!-- | H2b: Intraspecific competition | 1. specialization ~ 1/conspecific-packing              | -->
<!-- |                                | 2. trait-matching ~ 1/conspecific-packing              | -->
<!-- +--------------------------------+--------------------------------------------------------+ -->

<!-- **Table 1:** Hypothesized competitive processes with empirical predictions. BB = bumble bee; FL = floral. \ -->


<!-- In this context, we formulate a "metahypothesis" concerning the variation of hypothesized competition processes through elevation and time (**Figure 1**). Consistent with the stress-gradient hypothesis [@Bertness1994-zx] and with previous empirical work in our study system [@Hoiss2012-sn], we hypothesize an inverse relationship between competition and elevation. Along a temporal axis, we hypothesize that the strength of competition is structured by the seasonal life history pattern of bumble bee colonies [@Goulson2010-lo]. During the solitary phase, when the bumble bee community consists only of foundress queens, competition should be low. Both inter- and intraspecific competition should intensify as colonies increase in size during the social phase. After the switch point, when colonies begin producing gynes and males instead of workers, intraspecific comeptition should decrease due to the strong morphological differences between queens, males, and workers within each species, but interspecific competition might intensify due to the effective "over-packing" of trait space wih the introduction of multiple morphological castes per species.  -->

<!-- *Seasonal-Elevational metahypothesis.*    That any of the processes hypothesized above should obtain universally is both empirically and theoretically doubtful [@Ranta1981-vx]. More plausibly, which of these processes prevails -- and which species are favored by the prevailing process -- vary in both space and time, and this heterogenity of process could explain the structural idiosyncrasies of observed bumble bee communities [@Ranta1981-vx]. Rather than viewing this heterogeneity as an obstacle to inferring some universal rule of bumble bee community assembly, our aim in the present study is to investigate the heterogeneity itself as an ecological phenomenon, characterizing the variation in competitive processes (or lack thereof) in bumble bee communities through time and space.  -->

<!-- ```{r echo = FALSE, fig.cap = "**Figure 1:** Hypothesized variation in prevailing competitive processes through time and elevation."} -->
<!-- Time <- c(0,1,2,3) -->
<!-- Elevation <- c(0,1,2,3) -->
<!-- tab <- cbind(Time, Elevation) %>% -->
<!--   as_tibble() -->

<!-- ggplot(tab, aes(Time, Elevation)) + -->
<!--   geom_vline(xintercept = c(1,2), color = "white") + -->
<!--   geom_hline(yintercept = c(1,2), color = "white") + -->
<!--   scale_x_continuous(expand = c(0,0), -->
<!--                      breaks = c(0.5, 1.5, 2.5),  -->
<!--                      labels = c("founding", "build-up", "reproductive"), -->
<!--                      limits = c(0,3)) + -->
<!--   scale_y_continuous(expand = c(0,0), -->
<!--                      breaks = c(0.5, 1.5, 2.5),  -->
<!--                      labels = c("low", "mid", "high"), -->
<!--                      limits = c(0,3)) + -->
<!--   #geom_text() + -->
<!--   annotate("text", x = 0.5, y = 2.5, label = "H1") + -->
<!--   annotate("text", x = 1.5, y = 2.5, label = "H1") + -->
<!--   annotate("text", x = 2.5, y = 2.5, label = "H1") + -->
<!--   annotate("text", x = 0.5, y = 1.5, label = "H1") + -->
<!--   annotate("text", x = 0.5, y = 0.5, label = "H1") + -->
<!--   annotate("text", x = 1.5, y = 1.5, label = "H2 | H3") + -->
<!--   annotate("text", x = 2.5, y = 1.5, label = "H2 < H3") + -->
<!--   annotate("text", x = 1.5, y = 0.5, label = "H2 | H3") + -->
<!--   annotate("text", x = 2.5, y = 0.5, label = "H2 < H3") + -->
<!--   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) -->
<!-- ``` -->
<!-- \ -->



**References**

