---
title: "Beta Diversity through Elevation"
output: html_document
---

### Prepare environment
```{r message=FALSE, warning=FALSE}
library(sf)
library(gdm)
library(patchwork)
library(bipartite)
library(vegan)
library(ecodist)
library(gt)
library(tidyverse)
library(lubridate)
```

### Define functions
```{r}
# A function for converting a geometry column back to x and y columns
# https://github.com/r-spatial/sf/issues/231
sfc_as_cols <- function(x, names = c("x","y")) {
  stopifnot(inherits(x,"sf") && inherits(sf::st_geometry(x),"sfc_POINT"))
  ret <- sf::st_coordinates(x)
  ret <- tibble::as_tibble(ret)
  stopifnot(length(names) == ncol(ret))
  x <- x[ , !names(x) %in% names]
  ret <- setNames(ret,names)
  dplyr::bind_cols(x,ret)
}

# Function to convert networks to webs arrays
web_prep <- function(net) {
  
  out <- net %>%
    filter(caste != "male") %>%
    group_by(site, bb.sp, plant.sp) %>%
    summarize(freq = n()) %>%
    group_by(site) %>%
    mutate(bb.sp.count = length(unique(bb.sp)),
           plant.sp.count = length(unique(plant.sp))) %>%
    ungroup() %>%
    dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
    data.frame() %>% # convert to data frame
    frame2webs(type.out = "array")
}

# Function for betalinkr settings and conversion to tibble()
betalinkr_set1 <- function(x) {
  betalinkr_multi(x, partition.st = TRUE, partition.rr = TRUE, # partition both species turnover and richness/replacement
                  binary = TRUE, # treating beta-diversity as binary is an appropriate complement to our abundance analysis
                  index = "jaccard") %>% # the Jaccard index is best, I think, for binary analysis
  as_tibble() # for convenience
}

# Function for converting visitation data into site-pair table for GDM analysis 

interaction_gdm_prep <- function(x) {
  
  spdat <- net %>%
    filter(caste != "male") %>%
    dplyr::select(bb.sp, plant.sp.abb, site) %>% 
    distinct() %>% # each unique bb-fl-site combo
    mutate(present = rep(1, length(.[,1]))) %>% # we will treat abundance as binary
    unite(interaction, c(bb.sp, plant.sp.abb), sep = "_") %>% # unite bb and fl into interaction "species", i.e. a unique bb-fl combo
    pivot_wider(names_from = interaction, values_from = present) %>% # spread so that columns are interaction "species"
    replace(is.na(.), 0) %>% # replace NAs with absences
    as_tibble()

  envdat <- net %>%
    filter(caste != "male") %>%
    group_by(site) %>%
    summarize(samples = length(unique(date))) %>%
    left_join(site_data) %>% # add site data
    ungroup() %>%
    mutate(mowing = if_else(management == "mowing", 1, 0),
           grasing = if_else(management == "grasing", 1, 0),
           none = if_else(management == "none", 1, 0)) %>%
    dplyr::select(site, elev.mean, x, y, samples, mowing, grasing, none) %>%
    distinct() # return unique site-dates

  # Prep site-pairs for GDM analysis
  sitepair <- formatsitepair(spdat, 
                             bioFormat = 1, # i.e. samples as rows, species as columns, site-dates as rows
                             siteColumn = "site", 
                             dist = "jaccard", # I think Jaccard is the right choice
                             XColumn = "x", YColumn = "y", # spatial columns
                             predData = envdat, 
                             abundance = FALSE) %>% # abundance = FALSE means binary
    filter(s1.elev.mean != s2.elev.mean) # this drops within-site comparisons
}

# Function for preparing the floral survey biodata input to gdm's formatsitepair function
biodata_prep_survey <- function(surv) {
  surv %>%
    filter(flower.cover > 0) %>%
    mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
    group_by(site, plant.sp) %>%
    summarize(cover = sum(flower.cover)) %>%
    mutate(cover = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
    spread(plant.sp, cover) %>%
    replace(is.na(.), 0) %>%
    arrange(site)
}

# Function for preparing the floral survey predictor data input to gdm's formatsitepair function
predData_prep_survey <- function(surv) {
  surv %>%
    filter(flower.cover > 0) %>%
    group_by(site) %>%
    summarize(samples = length(unique(date))) %>%
    dplyr::select(site, samples) %>%
    left_join(site_data) %>%
    ungroup() %>%
    mutate(mowing = if_else(management == "mowing", 1, 0),
           grasing = if_else(management == "grasing", 1, 0),
           none = if_else(management == "none", 1, 0)) %>%
    dplyr::select(site, elev.mean, x, y, samples, mowing, grasing, none) %>%
    distinct() %>%
    arrange(site) 
}

# Function for preparing the visitation biodata input to gdm's formatsitepair function
biodata_prep_BB <- function(net) {
  net %>%
    filter(caste != "male") %>%
    group_by(site, bb.sp) %>%
    summarize(count = n()) %>%
    mutate(count = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
    spread(bb.sp, count) %>%
    replace(is.na(.), 0) %>%
    ungroup()
}

# Function for preparing the visitation predictor data input to gdm's formatsitepair function
predData_prep_BB <- function(net) {
  net %>%
    filter(caste != "male") %>%
    group_by(site) %>%
    summarize(samples = length(unique(date))) %>%
    dplyr::select(site, samples) %>%
    left_join(site_data) %>%
    ungroup() %>%
    mutate(mowing = if_else(management == "mowing", 1, 0),
           grasing = if_else(management == "grasing", 1, 0),
           none = if_else(management == "none", 1, 0)) %>%
    dplyr::select(site, elev.mean, x, y, samples, mowing, grasing, none) %>%
    unique() %>%
    arrange(site) 
}

# Function with setting for gdm's formatsitepair function
format_set <- function(bio, pred) {
  formatsitepair(bio, bioFormat = 1, siteColumn = "site", 
                 dist = "jaccard", XColumn = "x", YColumn = "y", 
                 predData = pred, abundance = FALSE) %>%
  filter(s1.elev.mean != s2.elev.mean) # this avoids comparing repeated measures of the same site (taking advantage of the fact that each site has a unique elevation in our dataset)
} 

# Function for tabulating gdm.varImp output 
gdm_tab <- function(x, response) {
  x %>%
    as.data.frame() %>%
    rownames_to_column(var = "predictor") %>%
    mutate(response = rep(response, n()),
           Predictor = case_when(
             predictor == "elev.mean" ~ "elevation",
             predictor == "samples" ~ "sampling freq.",
             predictor == "mowing" ~ "mowing",
             predictor == "grasing" ~ "grasing",
             predictor == "none" ~ "none")) %>%
    select(response, Predictor, Importance, pval) %>%
    tibble()
}
```

### Load site data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean, elev.class, elev.class2)) %>% #
  mutate(site = factor(site),
         tree_line = factor(tree_line)) %>% 
  semi_join(read_csv("../data/processed_data/network.csv"), by = "site") %>% # drop sites that were not sampled
  st_as_sf(coords = c("lon", "lat")) %>%
  st_set_crs(4326) %>%
  st_transform(5684) %>%
  sfc_as_cols() %>%
  as_tibble() %>%
  dplyr::select(-geometry)
```

### Load network and survey data
```{r}
net <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste  != "male") %>%
  unite(site.year.day, site, year, dayofyear, sep = "_", remove = FALSE) %>% 
  mutate(year = factor(year)) 

survey <- read_csv("../data/processed_data/floral_survey.csv") %>%
  filter(flower.cover > 0) %>%
  semi_join(net, by = "plant.sp") %>% # Consider only visited plant spp. (anytime, anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover) 
```

### Get beta-diversity for BBs and plants (not their interactions)
```{r}
# Bumble bee beta-diversity based on interaction sampling
bb_beta <- net %>%
  group_by(site, bb.sp) %>%
  summarize(count = n()) %>% 
  pivot_wider(names_from = bb.sp, values_from = count) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("site") %>%
  vegdist(method = "jaccard", binary = TRUE) %>% # Jaccard and binary settings make this comparable to the way I'm running betalinkr
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column(var = "site1") %>%
  pivot_longer(cols = -site1, names_to = "site2", values_to = "BB") 

# Flora beta-diversity based on floral survey data (filtered to include only species visited by BBs)
fl_beta <- survey %>%
  group_by(site, plant.sp) %>%
  summarize(count = sum(flower.cover)) %>%
  pivot_wider(names_from = plant.sp, values_from = count) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("site") %>%
  vegdist(method = "jaccard", binary = TRUE, upper = TRUE) %>%
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column(var = "site1") %>%
  pivot_longer(cols = -site1, names_to = "site2", values_to = "FL")

# Join into single data frame for visualization
species_beta <- full_join(bb_beta, fl_beta) %>%
  left_join(site_data, by = c("site1" = "site")) %>% # add site data
  dplyr::select(site1, site2, FL, BB, # first for site 1
                elev1 = elev.mean, x1 = x, y1 = y, mgmt1 = management) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, FL, BB, # then for site 2
                elev1, elev2 = elev.mean, x1, y1, x2 = x, y2 = y, mgmt1, mgmt2 = management) %>%
  mutate(site1 = factor(site1), 
         site2 = factor(site2),
         elev.diff = elev1 - elev2,
         mean.elev = (elev1 + elev2)/2, mgmt.diff = if_else(mgmt1 == mgmt2, 0, 1)) %>%
  pivot_longer(cols = c(BB, FL), names_to = "group", values_to = "beta") %>%
  filter(elev.diff >= 0) # For each site pair, one will be higher and the other lower; thus, when their order is reversed only the sign of the difference changes. Since no two sites have exactly the same elevation, this deduplicates the data by dropping the ordering of a given pair of sites (one of the two orderings) that yields a negative elevation difference while retaining the alternative ordering that yields a positive elevation difference. By >= 0, I retain comparisons of a site to itself. I drop (in the ggplot call) within-site comparisons when I plot against elevation difference, and I isolate within-site comparison when I plot against time. In the latter case, it is also necessary to deduplicate, since each pair of site dates occurs in A-B and B-A. For that, I filter (again, in the ggplot call) so that yday.diff > 0, similarly dropping the orderings that yield negative differences. Note that all this deduplication nonsense is not necessary for the betalinkr call because it deduplicates internally.


# fl_beta_dist <- survey %>%
#   group_by(site, plant.sp) %>%
#   summarize(count = sum(flower.cover)) %>%
#   pivot_wider(names_from = plant.sp, values_from = count) %>%
#   replace(is.na(.), 0) %>%
#   column_to_rownames("site") %>%
#   vegdist(method = "jaccard", binary = TRUE, upper = TRUE) 
# 
# bb_beta_dist <- net %>%
#   group_by(site, bb.sp) %>%
#   summarize(count = n()) %>% 
#   pivot_wider(names_from = bb.sp, values_from = count) %>%
#   replace(is.na(.), 0) %>%
#   column_to_rownames("site") %>%
#   vegdist(method = "jaccard", binary = TRUE)
```

### Tabulate webs
S	= beta_S, the dissimilarity in species composition
OS = beta_OS, the dissimilarity (component) explained by "rewiring" among shared species
WN = beta_WN, the dissimilarity between the two networks
ST = beta_ST, the dissimilarity (component) explained by difference in species community composition 
```{r message=FALSE}
# Create webs
net_web <- web_prep(net)

# Calculate beta-diversity
beta_site <- betalinkr_set1(net_web) %>%
  rename(site1 = i, site2 = j)

# Add site data
beta_mit_data <- beta_site %>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1 = elev.mean, x1 = x, y1 = y, mgmt1 = management) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2 = elev.mean, x1, y1, x2 = x, y2 = y, mgmt1, mgmt2 = management) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         mgmt.diff = if_else(mgmt1 == mgmt2, 0, 1)) %>%
  gather(metric, value, -c(site1, site2, elev1, elev2, 
                           elev.diff, mean.elev, 
                           x1, y1, x2, y2, mgmt1, mgmt2, mgmt.diff)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"),
    metric = factor(metric, levels = c("WN", "ST", "OS", "S", "ST.h", "ST.l", 
                                       "ST.lh", "WN.repl", "WN.rich", 
                                       "OS.repl", "OS.rich")))

# Write to csv for archiving
#write_csv(beta_mit_data, "../Dryad_ms1/partitioned_beta_diversity")
```

### Plot interaction beta-diversity against elevation difference
```{r}
# Set palette
p1 <- c("OS" = "#e7298a", "ST" = "#66a61e", "WN" = "black")

p2 <- c("ST.h" = "#7570b3", "ST.l" = "#1b9e77", "ST.lh" = "#d95f02")

# BB + FL elevation diff
g1 <- ggplot(filter(species_beta, 
                    site1 != site2), # no within-site comparisons
             aes(elev.diff, beta, color = group)) +
  geom_point(alpha = 0.25, pch = 16) +
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"), 
              se = FALSE) +
  xlab("Elevation difference (m)") +
  ylab("β-diversity (proportional species turnover)") +
  theme_light()

# BB * FL elevation diff (aggregate)
g2 <- ggplot(filter(beta_mit_data, 
                      site1 != site2 & # no within-site comparisons
                      metric != "S" &
                      metric.class == "aggregate"), 
             aes(elev.diff, value, color = metric))  +
  geom_point(alpha = 0.20, pch = 16) +
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"), 
              se = FALSE) +
  scale_color_manual(values = p1, name = "β partition") +
  xlab(NULL) +
  ylab("Prop. β-diversity") +
  theme_light()

# BB * FL elevation diff (ST partition)
g3 <- ggplot(filter(beta_mit_data, 
                      site1 != site2 & # no within-site comparisons
                      metric.class ==  "ST.partition"), 
             aes(elev.diff, value, color = metric))  +
  geom_point(alpha = 0.20, pch = 16) +
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"), 
              se = FALSE) +
  scale_color_manual(values = p2, name = "ST partition") +
  xlab("Elevation difference (m)") +
  ylab("Prop. species turnover") +
  theme_light()

# Patchwork plot for comparing species beta-diversity with interaction beta-diversity
png("../ms_1/figures/beta_stack.png", width = 6.5, height = 4.5, units = "in", res = 300)
(g1 | (g2 / g3)) + 
  plot_annotation(tag_levels = c("A", "B", "C")) + 
  theme_light() 
dev.off()

pdf("../ms_1/figures/beta_stack.pdf", width = 6.5, height = 4.5)
(g1 | (g2 / g3)) + 
  plot_annotation(tag_levels = c("A", "B", "C")) + 
  theme_light() 
dev.off()
```

### MRM analysis (not included in paper)
```{r, warning=FALSE}
survey_sample_diff <- survey %>%
  group_by(site) %>%
  summarize(samples = length(unique(date))) %>%
  remove_rownames() %>%
  column_to_rownames("site") %>%
  vegdist(method = "euclidean", upper = TRUE) %>%
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column(var = "site1") %>%
  pivot_longer(cols = -site1, names_to = "site2", values_to = "sample.diff")

net_sample_diff <- net %>%
  group_by(site) %>%
  summarize(samples = length(unique(date))) %>%
  remove_rownames() %>%
  column_to_rownames("site") %>%
  vegdist(method = "euclidean", upper = TRUE) %>%
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column(var = "site1") %>%
  pivot_longer(cols = -site1, names_to = "site2", values_to = "sample.diff")

mrm_survey_data <- species_beta %>%
  filter(group == "FL" & site1 != site2) %>%
  left_join(survey_sample_diff) %>%
  dplyr::select(site1, site2, beta, elev.diff, sample.diff, mgmt.diff) %>%
  mutate(elev.diff = scale(elev.diff), # gotta scale variables for MRM
         sample.diff = scale(sample.diff))

mrm_net_data <- species_beta %>%
  filter(group == "BB" & site1 != site2) %>%
  left_join(net_sample_diff)  %>%
  dplyr::select(site1, site2, beta, elev.diff, sample.diff, mgmt.diff) %>%
  mutate(elev.diff = scale(elev.diff),
         sample.diff = scale(sample.diff))

mrm_beta_data <- beta_mit_data %>%
  filter(site1 != site2) %>%
  left_join(net_sample_diff)  %>%
  dplyr::select(site1, site2, value, metric, elev.diff, sample.diff, mgmt.diff) %>%
  mutate(elev.diff = scale(elev.diff),
         sample.diff = scale(sample.diff))

# Function for tabulating modle coefficients
mrm_tab <- function(x, response) {
  x %>%
    as.data.frame() %>%
    rownames_to_column(var = "predictor") %>%
    slice(-1) %>%
    mutate(response = rep(response, n()),
           Predictor = case_when(
             predictor == "elev.diff" ~ "elevation",
             predictor == "sample.diff" ~ "sampling freq.",
             predictor == "mgmt.diff" ~ "management")) %>%
    select(response, Predictor, b, pval) %>%
    tibble()
}

# Call models
mrm_fl <- MRM(beta ~ elev.diff + sample.diff + mgmt.diff, 
              data = mrm_survey_data,
              nperm = 1000,
              method = "logistic")$coef %>%
  mrm_tab("Flora")

mrm_bb <- MRM(beta ~ elev.diff + sample.diff + mgmt.diff, 
              data = mrm_net_data,
              nperm = 1000,
              method = "logistic")$coef %>%
  mrm_tab("Bumble bees")

mrm_WN <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
              data = filter(mrm_beta_data, metric == "WN"),
              nperm = 1000,
              method = "logistic")$coef %>%
  mrm_tab("Total interactions")

mrm_ST <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
              data = filter(mrm_beta_data, metric == "ST"),
              nperm = 1000,
              method = "logistic")$coef %>%
  mrm_tab("ST partition")

mrm_ST.l <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
              data = filter(mrm_beta_data, metric == "ST.l"),
              nperm = 1000,
              method = "logistic")$coef %>%
  mrm_tab("ST.l partition")

mrm_ST.h <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
              data = filter(mrm_beta_data, metric == "ST.h"),
              nperm = 1000,
              method = "logistic")$coef %>%
  mrm_tab("ST.h partition")

mrm_ST.lh <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
              data = filter(mrm_beta_data, metric == "ST.lh"),
              nperm = 1000,
              method = "logistic")$coef %>%
  mrm_tab("ST.lh partition")

mrm_OS <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
              data = filter(mrm_beta_data, metric == "OS"),
              nperm = 1000,
              method = "logistic")$coef %>%
  mrm_tab("Rewiring partition")

# Tabulate results
mrm_results <- bind_rows(mrm_fl, mrm_bb, mrm_WN, mrm_ST, 
                         mrm_ST.h, mrm_ST.l, mrm_ST.lh, mrm_OS) %>%
  gt(groupname_col = "response") %>%
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = list(cells_column_labels(everything()))
  ) %>%
  tab_style(
    style = list(cell_text(style = "italic")),
    locations = list(cells_group(everything()))
  ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_body(
      columns = vars(pval),
      rows = pval < 0.05
    ))

gtsave(mrm_results,"../output/mrm_tab.pdf")

mrm_results_sub <- bind_rows(mrm_fl, mrm_bb, mrm_WN) %>%
  gt(groupname_col = "response") %>%
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = list(cells_column_labels(everything()))
  ) %>%
  tab_style(
    style = list(cell_text(style = "italic")),
    locations = list(cells_group(everything()))
  ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_body(
      columns = vars(pval),
      rows = pval < 0.05
    ))
```

### GDM: Interactions
```{r message=FALSE, warning=FALSE}
# Prep sitepairs
sitepair_WN <- interaction_gdm_prep(net)
sitepair_WNb <- interaction_gdm_prep2(net)

# Run GDM analysis
gdm_WN6 <- gdm(sitepair_WN, geo = FALSE, splines = c(6, 6, 6, 6, 6))
gdm_WN6b <- gdm(sitepair_WNb, geo = FALSE, splines = c(6, 6, 6, 6, 6))

gdm_WN6 <- gdm(sitepair_WN, geo = TRUE, splines = c(6, 6, 6, 6, 6, 6))


summary(gdm_WN)

gdm.varImp_WN <- gdm.varImp(sitepair_WN, geo = FALSE)
gdm.varImp_WN <- gdm.varImp(sitepair_WN, geo = FALSE, splines = c(6, 6, 6, 6, 6))
gdm.varImp_WN <- gdm.varImp(sitepair_WN, geo = FALSE, splines = c(6, 6, 6, 6))
gdm.varImp_WN <- gdm.varImp(sitepair_WN, geo = FALSE, splines = c(6, 6, 6))


# gdm.varImp_WN <- gdm.varImp(sitepair_WN, geo = FALSE, splines = c(6, 6, 6, 6))

# Plot
plot(gdm_WN6)
plot(gdm_WN6b)


# Plot with confidence intervals 
# pdf("../output/gdm_interactions.pdf", height = 2.5, width =  6.5)
# plotUncertainty(sitepair.WN, sampleSites = 0.7, bsIters = 100, geo  = FALSE, 
#                 splineCol = "black", plot.layout = c(1,3))
# dev.off()
```

### GDM: Floral survey
```{r}
# Prep input for GDM
bioData_survey <- biodata_prep_survey(survey)
predData_survey <- predData_prep_survey(survey)
sitepair_survey <- format_set(bioData_survey, predData_survey)

# Call GDMs
gdm_survey10 <- gdm(sitepair_survey, geo = TRUE, splines = c(10, 10, 10, 10, 10, 10))
gdm_survey6 <- gdm(sitepair_survey, geo = FALSE, splines = c(6, 6, 6, 6, 6))
gdm_survey5 <- gdm(sitepair_survey, geo = TRUE, splines = c(5, 5, 5, 5, 5, 5))
gdm_survey4 <- gdm(sitepair_survey, geo = TRUE, splines = c(4, 4, 4, 4, 4, 4))

summary(gdm_survey)
gdm.varImp_survey <- gdm.varImp(sitepair_survey, geo = FALSE, splines = c(6, 6, 6, 6, 6, 6))

# Plot
plot(gdm_survey10)
plot(gdm_survey6)
plot(gdm_survey5)
plot(gdm_survey4)


# Plot with confidence intervals (slow)
# pdf("../output/gdm_flora.pdf", height = 2.5, width =  6.5)
# plotUncertainty(sitepair_survey, sampleSites = 0.7, bsIters = 100, geo  = FALSE, 
#                 splineCol = "#00BFC4", plot.layout = c(1,3))
# dev.off()
```

### GDM: BB observations
```{r}
# Prep input for GDM
bioData_BB <- biodata_prep_BB(net)
predData_BB <- predData_prep_BB(net)
sitepair_BB <- format_set(bioData_BB, predData_BB)

# Call GDMs
gdm_BB6 <- gdm(sitepair_BB, geo = FALSE, splines = c(6, 6, 6, 6, 6))
gdm_BB5 <- gdm(sitepair_BB, geo = FALSE, splines = c(5, 5, 5, 5, 5, 5))
gdm_BB4 <- gdm(sitepair_BB, geo = TRUE, splines = c(4, 4, 4, 4, 4, 4))

summary(gdm_BB)
gdm.varImp_BB <- gdm.varImp(sitepair_BB, geo = TRUE, splines = c(6, 6, 6, 6, 6, 6))

# Plot
plot(gdm_BB6)
plot(gdm_BB5)
plot(gdm_BB4)


# Plot with confidence intervals (slow)
# pdf("../output/gdm_BB.pdf", height = 2.5, width =  6.5)
# plotUncertainty(sitepair_BB, sampleSites = 0.7, bsIters = 100, geo  = TRUE, 
#                 splineCol = "#F8766D", plot.layout = c(1,3))
# dev.off()
```

### Tabulate gdm.varImp results
```{r}
# Interaction model
gdm.varImp_WN.tab2 <- gdm.varImp_WN[[2]] %>%
  as.data.frame() %>%
  select(Importance = 1)

gdm.varImp_WN.tab3 <- gdm.varImp_WN[[3]] %>%
  as.data.frame() %>%
  select(pval = 1)

gdm.varImp_WN.tab <- bind_cols(gdm.varImp_WN.tab2, gdm.varImp_WN.tab3) %>%
  gdm_tab("Interactions") 

# Floral model
gdm.varImp_survey.tab2 <- gdm.varImp_survey[[2]] %>%
  as.data.frame() %>%
  select(Importance = 1)

gdm.varImp_survey.tab3 <- gdm.varImp_survey[[3]] %>%
  as.data.frame() %>%
  select(pval = 1)

gdm.varImp_survey.tab <- bind_cols(gdm.varImp_survey.tab2, gdm.varImp_survey.tab3) %>%
  gdm_tab("Flora") 

# BB model
gdm.varImp_BB.tab2 <- gdm.varImp_BB[[2]] %>%
  as.data.frame() %>%
  select(Importance = 1)

gdm.varImp_BB.tab3 <- gdm.varImp_BB[[3]] %>%
  as.data.frame() %>%
  select(pval = 1)

gdm.varImp_BB.tab <- bind_cols(gdm.varImp_BB.tab2, gdm.varImp_BB.tab3) %>%
  gdm_tab("Bumble bees") 

# Merge and format
gdm.varImp.tab <- bind_rows(gdm.varImp_WN.tab, gdm.varImp_survey.tab, gdm.varImp_BB.tab) %>%
  mutate(Importance = round(Importance, 2)) %>%
  gt(groupname_col = "response") %>%
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = list(cells_column_labels(everything()))
  ) %>%
  tab_style(
    style = list(cell_text(style = "italic")),
    locations = list(cells_row_groups(everything()))
  ) %>%
  data_color(
    columns = vars(Importance),
    colors = scales::col_numeric(
      palette = c("white", "red"),
      domain = c(-0, 56)), alpha = 0.75
    ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_body(
      columns = vars(pval),
      rows = pval < 0.05
    ))
  
gtsave(gdm.varImp.tab, "../output/gdm_tab.pdf")
```

### Fixed isplineExtract()
```{r}
##Extracts Ispline data from a gdm model
isplineExtract2 <- function (model){
  ###########################
  #model = gdmOb
  ###########################
  ##error checking
  ##checks to make sure a gdm model is given
  if(!is(model, "gdm")){
    stop("model argument must be of class = 'gdm'.")
  }
  
  ##Collects or sets simple data
  options(warn.FPU = FALSE)
  PSAMPLE <- 200
  preddata <- rep(0, times = PSAMPLE)
  pn <- model$predictors
  nPreds <- length(pn)
  yDat <- xDat <- matrix(0,PSAMPLE,nPreds)
  colnames(yDat) <- colnames(xDat) <- pn
  pmin <- 1
  pmax <- PSAMPLE
  
  ##cycles through each prodictor and fills the spline matrices
  splineindex <- 1
  for (i in 1:nPreds){ 
    #i<-1
    numsplines <- model$splines[i]
    z <- .C("GetPredictorPlotData", 
            pdata = as.double(preddata), 
            as.integer(PSAMPLE), 
            as.double(model$coefficients[splineindex:(splineindex + numsplines - 1)]), 
            as.double(model$knots[splineindex:(splineindex + numsplines - 1)]), 
            as.integer(numsplines),
            PACKAGE = "gdm")
    yDat[,i] <- z$pdata
    pmin <- pmin + PSAMPLE
    pmax <- pmax + PSAMPLE
    xDat[,i] <-  seq(from=model$knots[[(i*numsplines)-(numsplines-1)]],to=model$knots[[(i*numsplines)]], length=PSAMPLE)
    splineindex <- splineindex + numsplines
  }
  
  ##lists and returns matrices
  outData <- list(x=xDat,y=yDat)
  return(outData)
}
```

### Plot GDM splines
```{r}
# Interaction model
gdm_WN_splines1 <- isplineExtract2(gdm_WN6)$x %>%
  as.data.frame() %>%
  dplyr::select(elevation = elev.mean)

gdm_WN_splines2 <- isplineExtract2(gdm_WN6)$y %>%
  as.data.frame() %>%
  dplyr::select(response = elev.mean)

gdm_WN_splines <- bind_cols(gdm_WN_splines1, gdm_WN_splines2) %>%
  mutate(level = rep("Interactions"))

# Flora model
gdm_survey_splines1 <- isplineExtract2(gdm_survey6)$x %>%
  as.data.frame() %>%
  dplyr::select(elevation = elev.mean)

gdm_survey_splines2 <- isplineExtract2(gdm_survey6)$y %>%
  as.data.frame() %>%
  dplyr::select(response = elev.mean)

gdm_survey_splines <- bind_cols(gdm_survey_splines1, gdm_survey_splines2) %>%
  mutate(level = rep("Flora"))

# BB model
gdm_BB_splines1 <- isplineExtract2(gdm_BB6)$x %>%
  as.data.frame() %>%
  dplyr::select(elevation = elev.mean)

gdm_BB_splines2 <- isplineExtract2(gdm_BB6)$y %>%
  as.data.frame() %>%
  dplyr::select(response = elev.mean)

gdm_BB_splines <- bind_cols(gdm_BB_splines1, gdm_BB_splines2) %>%
  mutate(level = rep("Bumble bees"))

# Combine
gdm_splines <- bind_rows(gdm_WN_splines, gdm_survey_splines, gdm_BB_splines) %>%
  mutate(Level = factor(level, levels = c("Interactions", "Flora", "Bumble bees")))

tree_line <- site_data %>%
  dplyr::select(elevation = elev.mean, tree_line)

# Plot
ggplot(gdm_splines, aes(elevation, response, linetype = Level)) +
  geom_line() +
  #geom_vline(xintercept = 1500) +
  geom_rug(data = tree_line, aes(elevation, color = tree_line), inherit.aes = FALSE) +
  labs(color = "Tree Line", x = "Elevation", y = "Response") +
  theme_light(14) 

ggsave("../output/gdm_splines2.pdf")
ggsave("../output/gdm_splines2.png")

```

